04-analisis

### 2024-05-27 04-analisis.qmd
Update methods and analisis
update descriptive analisis market justice and meritocracy
update n lvl 2

---
author: "Equipo Edumer"
format: pdf
editor: visual
---

## Analysis

### Descriptive analysis

```{r packages, echo=FALSE}
pacman::p_load(tidyverse,
               ggpubr, 
               lme4, 
               ordinal,
               sjPlot, 
               texreg,
               kableExtra)

@@

@@

### Multivariate results

@@

```{r echo=FALSE}

@@

```{r echo=FALSE}
#| fig-cap: "Odds-ratios of justification of pensions, education and health by social and school meritocracy"
#| label: fig-odds
#| fig-asp: 0.8

plot_models(reg5_salud, reg5_educ, reg5_pension,  
            rm.terms = c("1|2", "2|3", "3|4", "educacion_recEd Media", "educacion_recEd Tecnica", "educacion_recUniversidad o posgrado", "educacion_recNs/Nr", "libros_recMás de 25", "acc_tec", "internet_rec2", "internet_recNs/Nr", "mean_educ", "cod_depe2Part. subvencionado", "cod_depe2Part. privado", "cod_grupoMedio bajo", "cod_grupoMedio", "cod_grupoMedio alto", "cod_grupoAlto", "simceMedio", "simceAlto"), #quitar terminos
            title = "", #quitar titulo
            show.values = TRUE, #mostrar valor de efectos
            dot.size = 1.3, #tamaño circulos
            line.size = 1, #tamaño CI
            value.size = 3.5, #tamaño valor efectoss
            spacing = .7, #espacio entre efectos
            grid = TRUE, # separa en un plot por modelo
            vline.color = "firebrick", # linea roja en punto neutro (1)
            axis.labels = rev(c("School talent", 
                            "School effort", 
                            "Social talent",
                            "Social effort", 
                            "Deservingness"
                            )), #con rev porque automatico los tira en otro orden
            show.legend = FALSE,
            m.labels = c("Health", "Education","Pension"),
            colors = gray(level = 0.2))  # variables dependientes

@@

pacman::p_load(kableExtra)
load("input/data/interact.RData")
kbl(collapse_rows,
    align = "lccccc", booktabs = T,
    col.names =  c("Variable", 
                   "School talent", 
                   "School effort",
                   "Social talent",
                   "Social effort", 
                   "Deservingness")) %>%
  add_header_above(c(" " = 1,
                     "Market Justice Preferences"=5), 
                   bold = T)%>%  
  kable_styling(latex_options = c("scale_down","HOLD_position"),
                full_width = F) %>%
  column_spec(1, bold = T, width = "4cm") %>%
  column_spec(column = 2:6,width = "1.5cm") %>%
  collapse_rows(columns = 1:2, valign = "middle") %>% 
  footnote(general = "***p < 0.001, **p < 0.01, *p < 0.05",
           footnote_as_chunk = T,general_title = "Note: ")

>>>>

---
author: "Equipo Edumer"
editor: visual
---

# Analysis

## Descriptive analysis

```{r packages, echo=FALSE}
pacman::p_load(tidyverse,
               ggpubr, 
               lme4, 
               ordinal,
               sjPlot, 
               texreg,
               kableExtra,
               sjPlot)

@@

*Market justice preferences*

```{r echo=FALSE}
#| fig-cap: "Market justice preferences in education, health and pensions"
#| fig-asp: 1.22
#| label: fig-marketjustice

data_rec %>% 
  select(just_salud, just_educ, just_pension) %>% 
  sjPlot::plot_stackfrq(expand.grid =TRUE,
                      geom.colors = "PuBu",
                      show.total = FALSE, 
                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
                      axis.labels = c("Health", "Education", "Pensions")) +
  theme_light(base_size = 16) +
  theme(legend.position="bottom", text = element_text(size = 14),
        plot.caption = element_text(hjust = 1))
```

@fig-marketjustice shows the frequency distribution for the three market justice preference items (pensions, education, and health). The majority of respondents disagree (63.0% in total) with the idea that it is just for higher-income individuals to access better pensions. However, a notable portion (27.0%) agrees, and a smaller minority (10.1%) strongly agrees. Similar to pensions, a majority of respondents (58.8%) disagree with the idea that higher-income individuals should have access to better education. However, a considerable portion (28.6%) agrees, and a slightly larger minority (12.6%) strongly agrees compared to pensions. In the context of health, an even larger majority of respondents (66.4%) disagree with the notion that higher-income individuals should access better healthcare services. A smaller percentage (23.6%) agrees, and a minority (10.0%) strongly agrees. Overall, @fig-marketjustice shows a predominant sentiment against the justice of higher-income individuals accessing better services across all three areas. The strongest opposition is observed in the health sector. Despite the overall disagreement, a significant portion of respondents agree with the statement, particularly in education, where agreement and strong agreement are slightly higher compared to pensions and health.

*School and social meritocracy*

```{r echo=FALSE}
#| fig-cap: "Social and school meritocracy"
#| fig-asp: 1.22
#| label: fig-meritocracy

plot<-data_rec %>% 
  select(merito_soc, esfuerzo_soc, inteligencia_soc, esfuerzo_esc, inteligencia_esc) %>% 
  sjPlot::plot_stackfrq(expand.grid =TRUE,
                      geom.colors = "PuBu",
                      show.total = FALSE, 
                      vjust=c(rep(c("center"),12), "bottom", "top", rep(c("center"), 6)),
                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
                      axis.labels = c("Deservingness", "Social effort", "Social talent", "School effort", "School talent")) +
  theme_light(base_size = 16) +
  theme(legend.position="bottom", text = element_text(size = 14),
        plot.caption = element_text(hjust = 1))

plot + 
  annotate("text", x = 5.5, y = 0, label = "School meritocracy", size = 5, hjust = 0, vjust = 1) +
  annotate("text", x = 3.5, y = 0, label = "Social meritocracy", size = 5, hjust = 0, vjust = 1)
```

@fig-meritocracy shows the frequency distribution for the five meritocracy items for the dimensions of school and social meritocracy. While a significant portion of respondents (29.2%) strongly agrees that talent is important to get good grades, a larger majority (78.6% combining agree and strongly agree) supports this notion. However, the strong agreement is not as pronounced as for school effort. An overwhelming majority (73.1%) agree that students effort is important to get good grades, which is a much higher percentage compared to the strong agreement for talent (29.2%). This indicates that people believe effort is more important than talent at school. There is a strong belief that both effort and talent are important, with a particularly high agreement for effort (73.1% strongly agree).

In relation with social meritocracy, @fig-meritocracy shows that a majority of respondents (63.7% combining agree and strongly agree) agree that people are rewarded for their talent in society, with a notable portion (17.1%) strongly agreeing. A majority (62.0% combining agree and strongly agree) agree that people are rewarded for their effort in society. Similar to social talent, the perception is positive but less overwhelming than school meritocracy. The perception of whether people get what they deserve in society is more divided. While a slim majority (56.5% combining agree and strongly agree) agree, a significant minority (43.6%) disagrees, showing that this is the most contentious item among respondents. There is general agreement that people are rewarded for talent and effort in society, but the consensus is weaker compared to school meritocracy. Deservingness in society is the most contentious, with significant disagreement among respondents.

*Market justice preferences in education, health and pensions by social and school meritocracy*

@@

## Models

@@

```{r echo=FALSE, warning=FALSE}

@@

```{r echo=FALSE, warning=FALSE}
#| fig-cap: "Odds-ratios of justification of pensions, education and health by social and school meritocracy"
#| label: fig-odds
#| fig-asp: 0.8

plot_models(reg5_salud, reg5_educ, reg5_pension,  
            rm.terms = c("1|2", "2|3", "3|4", "educacion_recEd Media", "educacion_recEd Tecnica", "educacion_recUniversidad o posgrado", "educacion_recNs/Nr", "libros_recMás de 25", "acc_tec", "internet_rec2", "internet_recNs/Nr", "mean_educ", "cod_depe2Part. subvencionado", "cod_depe2Part. privado", "cod_grupoMedio bajo", "cod_grupoMedio", "cod_grupoMedio alto", "cod_grupoAlto", "simceMedio", "simceAlto"), #quitar terminos
            title = "", #quitar titulo
            show.values = TRUE, #mostrar valor de efectos
            dot.size = 1.3, #tamaño circulos
            line.size = 1, #tamaño CI
            value.size = 3.5, #tamaño valor efectoss
            spacing = .7, #espacio entre efectos
            grid = TRUE, # separa en un plot por modelo
            vline.color = "firebrick", # linea roja en punto neutro (1)
            axis.labels = rev(c("School talent", 
                            "School effort", 
                            "Social talent",
                            "Social effort", 
                            "Deservingness"
                            )), #con rev porque automatico los tira en otro orden
            show.legend = FALSE,
            m.labels = c("Health", "Education","Pensions"),
            colors = gray(level = 0.2))  # variables dependientes

@@

pacman::p_load(kableExtra)
load("input/data/interact.RData")
kbl(collapse_rows,
    align = "lccccc", booktabs = T,
    col.names =  c("Variable", 
                   "School talent", 
                   "School effort",
                   "Social talent",
                   "Social effort", 
                   "Deservingness")) %>%
  add_header_above(c(" " = 1,
                     "Market Justice Preferences"=5), 
                   bold = T)%>%  
  kable_styling(latex_options = c("scale_down","HOLD_position"),
                full_width = F) %>%
  column_spec(1, bold = T, width = "4cm") %>%
  column_spec(column = 2:6,width = "1.5cm") %>%
  collapse_rows(columns = 1:2, valign = "middle") %>% 
  footnote(general = "*** p<0.001, ** p<0.01, * p<0.05",
           footnote_as_chunk = T,general_title = "Note: ")
